Design and Implementation of Probabilistic Programming Language Anglican

### 2017-03-14 paper/paper.tex

A probabilistic program, or query, mostly runs deterministic code.
Aside from the special forms \texttt{sample} and \texttt{observe},
which are probabilistic in nature, Anglican can be implemented as a
regular programming language. At \texttt{sample} and
\texttt{observe} forms, normal deterministic execution is 
interrupted and Anglican programs must allow the inference algorithm
to step in, recording information and affecting control flow. We refer these points in the execution as checkpoints. Handling of 
checkpoints can be implemented through coroutines/cooperative multitasking, and parallel execution/preemptive multitasking, as well as through explicit maintenance of program continuations. Anglican follows the latter option.

@@

The purpose of probabilistic forms \texttt{sample} and
\texttt{observe} is to interrupt deterministic computation and
transfer control to the inference algorithm.  Practically, this
is achieved by returning a record that represents the state of program execution at the checkpoint, rather than invoking a continuation. The Clojure record used to interrupt execution at 
\texttt{sample} and \texttt{observe} forms are \texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}sample} and \texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}observe} respectively. These records contain a unique identifier (which we will discuss in detail in section \ref{sec:addressing}), the arguments to the special form, the continuation, and the state.
\begin{lstlisting}[style=default]
=> (cps-of-expression '(sample dist) 'cont)
(->sample 'S1 dist cont $state)
=> (cps-of-expression '(observe dist v) 'cont)
(->observe 'O1 dist v cont $state)
\end{lstlisting}
Here \texttt{->sample} and \texttt{->observe} in the
CPS-transformed expressions are constructors for Clojure
records, and they take values of their fields as arguments.
Calling the continuation resumes the computation.

>>>>

A probabilistic program, or query, mostly runs deterministic code.
Aside from the special forms \texttt{sample} and \texttt{observe},
which are probabilistic in nature, Anglican can be implemented as a
regular programming language. At \texttt{sample} and
\texttt{observe} forms, normal deterministic execution is 
interrupted, and Anglican programs must allow the inference algorithm
to step in, recording information and affecting control flow. We refer these points in the execution as checkpoints. Handling of 
checkpoints can be implemented through coroutines/cooperative multitasking, and parallel execution/preemptive multitasking, as well as through explicit maintenance of program continuations. Anglican follows the latter option.

@@

The purpose of probabilistic forms \texttt{sample} and
\texttt{observe} is to interrupt deterministic computation and
transfer control to the inference algorithm.  Practically, this
is achieved by returning a record that represents the state of program execution at the checkpoint, rather than invoking a continuation. The Clojure record used to interrupt execution at 
\texttt{sample} and \texttt{observe} forms are \texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}sample} and \texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}observe} respectively. These records contain a unique identifier (which we will discuss in detail in section \ref{sec:addressing}), the arguments to the special form, the continuation, and the state.
\begin{lstlisting}[style=default]
=> (cps-of-expression '(sample dist) cont)
(->sample 'S1 dist cont $state)
=> (cps-of-expression '(observe dist v) cont)
(->observe 'O1 dist v cont $state)
\end{lstlisting}
Here \texttt{->sample} and \texttt{->observe} in the
CPS-transformed expressions are constructors for Clojure
records, and they take values of their fields as arguments.
Calling the continuation resumes the computation.
